% embodimetry --- 4-5 page arXiv writeup (cs.RO primary, cs.LG secondary).
% Capability-Ladder Audit framing: L0 pretrained / L1 fine-tune / L2 classical
% / L3 world-model planner on ONE seed+Wilson-CI+MDE contract. The LEAD
% contribution is the auditable instrument + the self-caught normalization
% bug (ACT x aloha 0.016 -> 0.824 via a clean 2x2 ablation). L3 is an
% IN-FLIGHT EXISTENCE-PROOF / HYPOTHESIS (a dynamics-complexity gradient for
% WHEN zero-training planning substitutes for learning), NOT a confirmed
% result: the Wall number is small-N with a wide Wilson CI that spans chance,
% and the Wall-vs-PushT cross-env contrast is CONFOUNDED (env, checkpoint,
% and CEM budget co-vary). L3 numbers are a CROSS-REPO artifact
% (lerobot-wm-research), cited as such and demoted in certainty accordingly.

\documentclass[10pt]{article}

\usepackage[a4paper,top=0.7in,bottom=0.7in,left=0.75in,right=0.75in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{url}
\usepackage[compact]{titlesec}
\titlespacing*{\section}{0pt}{1.1ex plus 0.3ex minus 0.2ex}{0.6ex plus 0.2ex}
\titlespacing*{\subsection}{0pt}{0.8ex plus 0.2ex minus 0.2ex}{0.4ex plus 0.1ex}
\titlespacing*{\paragraph}{0pt}{0.6ex plus 0.2ex minus 0.1ex}{1em}

\hypersetup{
    colorlinks=true,
    linkcolor=blue!50!black,
    citecolor=blue!50!black,
    urlcolor=blue!50!black,
}

% \todo marks a number not yet pinned to an in-tree parquet/JSON value.
% Render it visibly so a human resolves it before submission.
\newcommand{\todo}[1]{\textcolor{red}{[\textbf{TODO:} #1]}}

\title{\vspace{-1.5ex}embodimetry: a Capability-Ladder Audit ---
measuring pretrained, fine-tuned, and classical robot policies on one
reproducible eval contract\vspace{-0.5ex}}
\author{Th\'eo Hermann\\
\texttt{thermann.ai@gmail.com} \quad
\texttt{https://github.com/thrmnn/embodimetry}}
\date{\vspace{-2ex}\today\vspace{-1ex}}

\begin{document}

\maketitle

% ---------------------------------------------------------------- %
% Abstract                                                         %
% ---------------------------------------------------------------- %

\begin{abstract}
The animating question of this work is an \emph{in-flight hypothesis} we
pose but do not answer: does manipulation success degrade along a
\emph{dynamics-complexity gradient}, with a zero-training world-model
planner versus a learned policy as the contrast --- from point-mass
navigation, through contact, to deformable manipulation? Answering it
requires a ruler that scores incompatible paradigms together, and
\emph{building that ruler is the citable contribution of this paper}; the
gradient itself is the first sharp question it is purpose-built to test,
not a result we claim here. Robot-policy success rates are reported per
paradigm --- pretrained imitation, fine-tuning, classical control,
world-model planning --- on incompatible protocols, so the paradigms
cannot be put on one axis. We introduce \emph{embodimetry}, a
\emph{Capability-Ladder Audit}, and stand behind two settled,
reproducible contributions. \emph{First, a single cross-paradigm
evaluation contract}: pretrained, fine-tuned, classical-control, and
world-model-planner policies are all scored as the \emph{same}
\texttt{PolicyCallable} under one seed / Wilson-and-bootstrap-CI /
minimum-detectable-effect (MDE) ruler on shared
LeRobot~\cite{cadene2024lerobot} tasks, with floor baselines and honest
negatives (a SmolVLA-LoRA collapse, a classical controller at $0.012$)
kept on the board --- no neighboring effort puts paradigms on one ruler
with the negatives retained. \emph{Second, a self-auditing methodology}:
the instrument caught and corrected its \emph{own} headline overclaim ---
a normalization bug in our harness pinned ACT on Aloha-transfer-cube at
$\hat{p}{=}0.016$ (below the env's random floor), and a clean $2{\times}2$
ablation isolated the full $0.016\to0.824$ recovery to the normalization
fix, not inference settings --- and a guard test code-enforces that the
unproven top rung stays labeled in-flight. Our claim is that an eval suite
must be audited as carefully as the policies it scores. The ladder is the
honest-negatives spine that earns the contract credibility. \textbf{L0}
(pretrained Diffusion Policy on PushT) scores $0.816$~[0.739, 0.874].
\textbf{L1} (continued fine-tuning of an already-converged ACT on
Aloha-transfer-cube) moves $0.824$~[0.772, 0.866] to $0.864$~[0.816,
0.901] --- a $+0.040$ shift whose CIs \emph{overlap} (Cohen's $h{=}0.11$,
below the $N{=}250$ MDE), reported as within noise; and an attempted
SmolVLA LoRA fine-tune collapses ($0.252\to{\sim}0$). \textbf{L2} (a
competent scripted PushT controller) clears the strict bar on only
$0.012$~[0.004, 0.035] of episodes despite mean rollout coverage near
$0.50$. On the gradient we report only an in-flight signal, never a
finding: the \emph{same} zero-training world-model planner
(DINO-WM~\cite{zhou2024dinowm} latent MPC) reaches Wall-nav at
$9/24{=}0.375$~[0.21, 0.57] while PushT (${\sim}0$) is \emph{the
contact endpoint of the hypothesized curve}, not a failure. This signal is
$N{=}24$, a single world-model family, and \emph{cross-env confounded}
(env, checkpoint, and CEM budget co-vary) --- we state that confound as an
open limitation, run the L3 cells as a cross-repo artifact
(\texttt{lerobot-wm-research}), and a second-WM-family de-confound
(\texttt{jepa-wms}) is in progress, so the gradient earns its own citation
only in v1.1. The contributions we stand behind without qualification are
the contract and the self-audit; the gradient is the in-flight hypothesis
the contract was purpose-built to measure.
\end{abstract}

% ---------------------------------------------------------------- %
% Introduction                                                     %
% ---------------------------------------------------------------- %

\section{Introduction}

The question that animates this work is one we \emph{pose}, not answer:
does manipulation success degrade along a \emph{dynamics-complexity
gradient} --- from point-mass navigation, through contact, to deformable
manipulation --- with a zero-training world-model planner versus a learned
policy as the contrast? We flag this up front as an \emph{in-flight
hypothesis}: the cleanest way to ask it is to score a world-model planner
and a learned policy on \emph{one} ruler, and \textbf{building that ruler,
not answering the gradient, is what this paper contributes}. We stand
behind two settled, reproducible contributions, and treat the gradient as
the first sharp question the ruler is purpose-built to test
(\S\ref{sec:l3}). \emph{(1) A single cross-paradigm evaluation contract.}
Robot control is pursued along a ladder of paradigms with increasing
autonomy and decreasing supervision: \textbf{L0} zero-shot evaluation of
a pretrained policy, \textbf{L1} fine-tuning that policy on a small demo
budget, \textbf{L2} hand-engineered classical control, and \textbf{L3}
planning through a learned world model. Each rung has its own literature
and its own way of reporting success, so a Diffusion Policy
number~\cite{chi2023diffusion}, an ACT fine-tuning
number~\cite{zhao2023learning}, a scripted-controller baseline, and a
world-model planner~\cite{zhou2024dinowm,assran2025vjepa2} are never
measured on a common axis. The protocols differ in seed counts, episode
budgets, success thresholds, and whether confidence intervals are
reported at all --- a reproducibility gap documented repeatedly in
adjacent fields~\cite{henderson2018deep,agarwal2021deep}. embodimetry
scores all four paradigms as the same \texttt{PolicyCallable} under one
seed / Wilson-and-bootstrap-CI / MDE ruler, with floor baselines and
honest negatives (the SmolVLA-LoRA collapse, the classical $0.012$) kept
on the board --- no neighbor (HF leaderboards, MLPerf, Open
X-Embodiment~\cite{bohg2025open}, or the world-model
benchmarks~\cite{zhou2024dinowm,assran2025vjepa2}) puts paradigms on one
ruler with the negatives retained. \emph{(2) A self-auditing methodology.}
The instrument caught and corrected its own headline overclaim, code-fenced
against re-inflation by a guard test (\S\ref{sec:methods}); we develop this
below as the load-bearing claim that an eval suite must be audited as
carefully as the policies it scores.

\emph{embodimetry} is a \emph{Capability-Ladder Audit}: it measures
multiple paradigms as the \emph{same} \texttt{PolicyCallable}
(\texttt{obs} $\to$ \texttt{action}) through \emph{one} seed /
Wilson-CI / MDE contract on shared LeRobot~\cite{cadene2024lerobot}
tasks. A pretrained Diffusion Policy, a fine-tuned ACT, and a scripted
PushT controller all enter the same eval loop, are seeded identically,
and are scored with the same Wilson interval, so their numbers sit on
one yardstick instead of in four PDFs. The instrument is paradigm-neutral
by construction: it has no stake in which rung wins.

We release embodimetry across three surfaces: open data (a public
Hugging Face Hub~\cite{huggingfacehub2024} dataset of every per-episode
outcome plus an MP4 of every rollout); open eval (a pinned, scriptable
pipeline that re-derives any parquet row in $<1$\,h on commodity
hardware); and open analysis (the notebook
\texttt{notebooks/01-write-finding.ipynb} that produces every table and
figure here). The v1 leaderboard is four learned policies and two
weights-free floor baselines crossed with six envs (22~runnable cells
after the checkpoint-availability filter); the ladder rungs (L1, L2) are
gated \emph{out} of that leaderboard and run as separate audit cells on
the identical contract.

The result we lead with is that the instrument audits itself. The
contract flagged a normalization bug in \emph{our own} eval harness ---
ACT on \texttt{aloha\_transfer\_cube} read $\hat{p}{=}0.016$, below the
env's random floor, until we traced it to un-normalized image
observations; a clean $2{\times}2$ ablation pins the entire
$0.016\to0.824$ recovery on the normalization fix, with temporal
ensembling a wash (\S\ref{sec:results-audit}). A loss curve would never
have surfaced this; only a closed-loop success rate with a confidence
interval did. That is the load-bearing claim of the paper: the value of a
neutral contract is that it bites on the surface it is supposed to
protect, including the auditor's own.

This paper reports all four rungs on the same contract. The L0--L2 rungs
are a spine of honest negatives that is what makes the contract
trustworthy: fine-tuning a near-ceiling policy moves the number only
\emph{within noise} (L1, ACT); an attempted SmolVLA LoRA fine-tune
collapses rather than lifts; and a competent classical controller clears
the strict success bar on ${\sim}1\%$ of episodes despite getting halfway
there on average (L2) --- a quantitative case that the value of learning
is the last fraction of precision.

At the top of the ladder we \emph{pose} an \emph{in-flight hypothesis},
not a result. We ask \emph{when} a single zero-training world-model
planner --- DINO-WM~\cite{zhou2024dinowm} latent MPC, no task-specific
training --- can stand in for learning, and conjecture a
\emph{dynamics-complexity gradient}: from point-mass navigation, through
contact, to deformable manipulation. The signal so far is an
\emph{in-flight} reading at the two ends, not a measured curve, and every
number below travels with its CI, its single-family caveat, and its
cross-env confound in the same breath. On PushT, contact-rich
manipulation, the planner scores ${\sim}0$ ($0/2$) --- \emph{the contact
endpoint of the hypothesized curve}, a clean data point where a frozen
latent model cannot plan, treated as a feature of the study rather than a
failure to rescue with compute. On Wall, a two-room point-mass navigation
task, the same planner reaches $9/24{=}0.375$ with a Wilson interval
[0.212, 0.573] wide enough to span chance (\S\ref{sec:l3}), on a
\emph{single} DINO-WM family. Crucially, the Wall-vs-PushT contrast is
\emph{confounded} --- env, checkpoint, and CEM budget all co-vary --- so it
is an in-flight signal that motivates a de-confounded study (same planner,
vary only the env), not a conditional we can claim; a second-WM-family
de-confound (\texttt{jepa-wms}) is in progress but throughput-blocked. The
contract is built precisely to make that study possible on one shared
seed/CI protocol; we report the hypothesis, the single-family limit, and
the open confound honestly rather than inflate indicative cells into a
finding. The contract tells you uncomfortable truths --- including about
its own claims; that is the point. The benchmark itself is the data; this
paper is the defense.

% ---------------------------------------------------------------- %
% Related Work                                                     %
% ---------------------------------------------------------------- %

\section{Related work}
\label{sec:related}

\paragraph{Robot-learning benchmarks and statistical rigor.}
PushT~\cite{chi2023diffusion}, the Aloha suite~\cite{zhao2023learning},
and LIBERO~\cite{liu2023libero} are the de facto LeRobot references;
RLBench~\cite{james2020rlbench}, CALVIN~\cite{mees2022calvin}, and Open
X-Embodiment~\cite{bohg2025open} offer larger taxonomies but no
pretrained LeRobot-compatible checkpoints or shared protocol. None ship a
multi-policy leaderboard with shared seed accounting, multi-seed CIs, and
re-runnable rows. We adopt the multi-seed, bootstrap-CI, MDE-bounded
posture that Henderson et al.~\cite{henderson2018deep} and Agarwal et
al.~\cite{agarwal2021deep} established for deep RL, and add weights-free
no-op/uniform-random floor baselines on every cell (per the
Gym~\cite{brockman2016openai} convention) since some envs admit non-zero
random success.

\paragraph{Cross-paradigm and world-model planning.}
The ladder we audit spans paradigms each benchmarked in isolation:
classical control and learned policies are compared only informally, and
world-model planners --- PlaNet~\cite{hafner2019planet},
DINO-WM~\cite{zhou2024dinowm}, V-JEPA-2-AC~\cite{assran2025vjepa2} ---
report success on their own tasks, never against a learned imitation
policy under one shared seed/CI protocol. Our contribution is not a new
policy class but the \emph{contract} that puts a pretrained, fine-tuned,
classical, \emph{and} world-model planner on the same Wilson interval ---
with floor baselines and honest negatives retained, which no neighboring
benchmark does --- plus the \emph{de-confounded cross-env study it makes
possible}: running the same DINO-WM planner on envs of differing dynamical
complexity \emph{to ask}, as an open question, \emph{when} zero-training
planning substitutes for learning. We \emph{pose} that study as an
in-flight hypothesis whose signal so far is confounded and single-family
(\S\ref{sec:l3}), not a measured conditional.

% ---------------------------------------------------------------- %
% Methods                                                          %
% ---------------------------------------------------------------- %

\section{Methods}
\label{sec:methods}

\paragraph{Sweep contract.}
The unit of evaluation is a \textit{cell}:
$(\text{policy}, \text{env}, \text{seed\_idx})$ with
$\text{seed\_idx} \in \{0,\ldots,4\}$. Each cell runs $n_\text{ep}$
episodes; the contract is $n_\text{ep}{=}50$ unless the auto-downscope
rule (below) reduces it, giving $N = 5 \cdot 50 = 250$ binary outcomes.
An episode runs from \texttt{env.reset(seed=$s$)} until
\texttt{terminated}, \texttt{truncated}, or step count reaches
env-specific \texttt{max\_steps} (PushT 300, Aloha 400, LIBERO
$\{280,280,300,520\}$ for spatial/object/goal/10, inherited from
lerobot defaults; canonical LIBERO uses 600 for all four, an audit
caveat in Section~\ref{sec:results-audit}). Success is the per-env
\texttt{SUCCESS\_REWARD} threshold; the episode succeeds iff its
final-step reward meets or exceeds it. PR~\#90 adds a selectable
\texttt{--canonical} criterion (paper termination rule, 600-step LIBERO
cap); v1 ships the lerobot-default rule and v1.1 reruns affected cells.

\paragraph{Seeding contract.}
Per cell, three RNGs are seeded with
$\text{seed}_\text{cell} = \text{seed\_idx} \cdot 1000$
(\texttt{numpy}, \texttt{torch}, \texttt{torch.cuda}). Per episode the
env is reset with $\text{seed}_\text{ep}(e) = \text{seed}_\text{cell}+e$.
Policy stochasticity inherits the torch generator and is \emph{not}
re-seeded per episode, so the generator advances across episodes within
a cell: mid-cell resume is \emph{not} bit-reproducible and a crashed
cell restarts from $e{=}0$. The orchestrator
(\texttt{checkpointing.plan\_resume}) classifies cells as completed,
partial, or pending and re-runs partial cells from scratch.

\paragraph{Confidence intervals and paired comparisons.}
Each cell reports $\hat{p}{=}\text{successes}/N$ with two mutually
sanity-checked 95\% CIs: the closed-form Wilson score
interval~\cite{wilson1927probable,agresti1998approximate} (shown in the
tables) and a percentile bootstrap~\cite{efron1979bootstrap,efron1993introduction}
($10{,}000$ resamples, seeded \texttt{Generator(42)}, resampling over
\emph{episodes} not seeds); the two agree within 1\,pp at $N{=}250$. For
two cells sharing $(\text{seed\_idx}, e)$ pairings we report a paired
$\Delta\hat{p}$ with a pivotal bootstrap CI and the Wilcoxon
signed-rank test~\cite{wilcoxon1945individual} (reducing to
McNemar~\cite{mcnemar1947note} for binary outcomes), plus Cohen's
$h{=}2\arcsin\!\sqrt{p_1}-2\arcsin\!\sqrt{p_2}$~\cite{cohen1988statistical}
($|h|{<}0.2$ small). Multiple within-env ranking claims are controlled by
Holm--Bonferroni~\cite{bonferroni1936teoria} at $\alpha{=}0.05$.

\paragraph{Minimum detectable effect.}
At $N{=}250$ the Wilson 95\% half-width at worst case $\hat{p}{=}0.5$ is
$\text{HW}=0.0615$; any $|\Delta\hat{p}| < 2\cdot\text{HW} = 0.1230$ is
inconclusive at this $N$ regardless of $p$-value, since two cells with
half-widths up to 0.0615 can produce a smaller sample delta from
sampling noise alone. The threshold shrinks for saturated/near-zero
cells ($0.0543$ at $\hat{p}{=}0.95$); \texttt{docs/MDE\_TABLE.md}
tabulates the per-$\hat{p}$ threshold and the empirical paired bootstrap
MDE at $\rho\in\{0,0.3,0.5\}$. Every paired comparison in
Section~\ref{sec:results} is annotated against this threshold; deltas
below it are reported as \emph{inconclusive at $N{=}250$}, not as
findings.

\paragraph{Auto-downscope rule.}
\label{sec:downscope}
The per-cell wall-clock budget is 3~hours on a single RTX~4060 laptop
GPU (8\,GB VRAM). Given a per-policy mean step latency $\overline{t}$
(ms/step) from Day-0 calibration, the episodes-per-seed count is
\begin{equation}
    n_\text{ep} = \min\!\left(50,\ \left\lfloor
        \frac{3 \cdot 3600}{5 \cdot \texttt{env.max\_steps} \cdot
        \overline{t} / 1000} \right\rfloor\right).
\end{equation}
Seed count is fixed at five; only $n_\text{ep}$ shrinks, so a
down-scoped cell lands at $N\in\{50,100\}$ with a wider CI flagged in
the caption. Two v1 cells were down-scoped. If the per-cell
minimum ($N{=}50$) still exceeds budget, the policy is dropped from v1,
not silently truncated.

\paragraph{Sparse matrix and reproducibility.}
A cell is N/A only when no public Hub checkpoint exists at lock-in,
reducing the $6\times6$ grid to 22~runnable cells; sparsity is a visible
gap, not silently skipped. Each parquet row carries its code SHA, the
lerobot pin (\texttt{0.5.1}), and the SHA-256 of its MP4; every policy is
locked to a fixed \texttt{revision\_sha} and the eval loop refuses to load
one without it. A single row replays via \texttt{run\_one.py}. The full
sweep took $\approx 25$\,h of GPU time on the reference hardware
(1$\times$ RTX 4060 Laptop, WSL2).

\paragraph{The same contract across rungs.}
\label{sec:ladder-contract}
Every rung enters the audit as a \texttt{PolicyCallable}: a function
mapping an observation dict to an action, with no knowledge of which
paradigm produced it. The L1 fine-tuned ACT, the L2 scripted controller,
and the L0 pretrained checkpoints all run through the \emph{identical}
\texttt{embodimetry.eval.run\_cell} loop --- the same per-cell seeding
contract, the same $5{\times}50$ budget, and the same env-canonical
success rule the corresponding zero-shot cell used. The only thing that
varies between an L0 and L1 cell is the policy weights; between an L0 and
L2 cell, the policy is weights-free scripted control. Ladder cells are
gated out of \texttt{V1\_POLICIES} (the published leaderboard roster) and
written to a separate \texttt{results/ladder/} tree, so an audit rung can
never leak into the leaderboard or rewrite a canonical parquet row, while
still being directly CI-comparable to it.

\paragraph{L1 rung --- continued fine-tuning.}
We full-fine-tune the pretrained ACT checkpoint on a small demo budget
from \texttt{aloha\_transfer\_cube} and re-measure on the contract its
zero-shot cell used (canonical \texttt{sticky\_reward\_eq}, reward${=}4$,
$N{=}250$). ACT (${\sim}50$\,M params) full-FT peaks at ${\sim}2.8$\,GB
VRAM and fits the 8\,GB budget; SmolVLA full-FT does not, so L1 is a
single-cell ACT probe (see Limitations). Fine-tuning an already-converged
policy on its own data is expected to yield a near-zero lift; we report
the lift with its paired effect size and call it within noise when its CI
straddles zero, rather than dropping the negative result.

\paragraph{L2 rung --- classical-control quality.}
The L2 cell runs a deterministic scripted state-feedback PushT controller
through the same loop under the strict canonical PushT rule
(\texttt{sticky\_is\_success}, $\text{coverage}>0.95$). Because a binary
success rate alone hides \emph{how close} a competent controller gets, we
additionally log the per-episode max-coverage
$\max_t \text{info}[\text{coverage}]_t$ (summary JSON only, never the
parquet schema): the gap between mean max-coverage and the strict $0.95$
bar quantifies the last-mile precision that the binary rate discards.

\paragraph{L3 rung --- zero-training world-model planning.}
The L3 cell is a \emph{zero-training} planner: a frozen
DINO-WM~\cite{zhou2024dinowm} latent forward model wrapped in a
cross-entropy-method (CEM) MPC loop, entering the audit as a
\texttt{PolicyCallable} with no task-specific training and its per-step
sampling cost logged alongside its success rate. At each step the planner
encodes the current observation once, rolls candidate action sequences
through the latent model, and returns the first action of the
lowest-cost sequence (cost ${=}$ latent distance to the encoded goal).
The \emph{same} planner and CEM hyperparameters are run on two envs of
differing dynamical complexity: \textbf{Wall}, a two-room point-mass
navigation task (success $:= \lVert\text{goal}-\text{pos}\rVert < 4.5$,
\texttt{max\_steps}${=}40$), and \textbf{PushT}, contact-rich
manipulation. The CEM budget is $N{=}300$ samples, $10$ optimization
iterations, top-$k{=}30$ elites, horizon $5$ --- a \emph{reduced} budget
chosen so the loop fits an 8\,GB GPU. A $39\times$ predictor speedup
(scaled-dot-product attention, \texttt{bfloat16}, and encode-once frame
caching) is what made any L3 measurement feasible on this hardware; even
so, the budget is well below the $300\times30$ CEM the source method
uses at H100 scale, which we flag as a magnitude (not direction)
caveat. These L3 cells live in a separate research repository
(\texttt{lerobot-wm-research}); we report them as a \emph{cross-repo}
artifact run on the same eval contract, and mark their figures' provenance
accordingly rather than fabricating an in-tree artifact. We stress that
the Wall and PushT cells differ in \emph{three} co-varying dimensions ---
env, checkpoint, and CEM budget --- so they constitute an in-flight signal
for a dynamics-complexity gradient, not a controlled conditional; the
de-confounded version varies only the env (\S\ref{sec:l3}).

\paragraph{Anti-inflation guard (the self-audit, code-enforced).}
\label{sec:guard}
The self-auditing methodology is not just a posture, it is a test.
\texttt{tests/test\_no\_ungated\_l3\_claims.py} runs in CI and \emph{fails
the build} if any public surface (this paper, the README, the living
leaderboard, the deck, the blog) asserts the world-model rung as a settled
result: it greps for ``substitute(s) for learning'' phrasings and requires
each to sit within a word window of a gating term (\emph{hypothesis},
\emph{in-flight}, \emph{existence-proof}, \emph{confound}), forbids the
specific triumphalist tokens the original audit flagged, and asserts the
$0.016\to0.824$ self-caught-bug lead is present on every surface. The
guard is citable evidence of the self-audit: it makes the in-flight status
of the L3 gradient a \emph{property of the repository}, not an editorial
promise, so the dynamics-complexity hypothesis cannot be silently
re-inflated into a claim between releases.

% ---------------------------------------------------------------- %
% Results                                                          %
% ---------------------------------------------------------------- %

\section{Results}
\label{sec:results}

\paragraph{The ladder on one contract.}
Table~\ref{tab:ladder} is the spine of the paper: the L0--L2
rungs --- pretrained, fine-tuned, classical --- read off the same Wilson
interval. The reading is deliberately undramatic. Fine-tuning a
near-ceiling policy moves it within noise; a competent scripted
controller gets halfway to the goal on average but almost never clears
the strict bar. Both statements are quantitative because all the rungs
share one statistical spine. The L3 rung (\S\ref{sec:l3}) sits on the
same contract as an in-flight hypothesis --- the same world-model planner
read across two envs --- and is reported as an existence-proof with a
named confound, not a measured result.

\begin{table}[t]
\centering
\small
\begin{tabular}{llrll}
\toprule
Rung & Policy $\times$ env & $N$ & $\hat{p}$ (Wilson 95\%) & Note \\
\midrule
L0 & Diffusion $\times$ pusht$^\dagger$ & 125 & 0.816 [0.739, 0.874] & pretrained zero-shot \\
L1 & ACT$_\text{zero-shot}$ $\times$ aloha & 250 & 0.824 [0.772, 0.866] & baseline \\
L1 & ACT$_\text{fine-tuned}$ $\times$ aloha & 250 & 0.864 [0.816, 0.901] & $\Delta{=}{+}0.040$, CIs overlap \\
L2 & classical $\times$ pusht & 250 & 0.012 [0.004, 0.035] & mean cov.\ ${\approx}0.50$ \\
\midrule
L3 & DINO-WM planner $\times$ wall & 24 & 0.375 [0.21, 0.57] ($9/24$) & zero-training nav (Table~\ref{tab:l3}) \\
L3 & DINO-WM planner $\times$ pusht & 2 & ${\sim}0.0$ ($0/2$) & same planner, contact endpoint \\
\bottomrule
\end{tabular}
\caption{The Capability-Ladder Audit: paradigms scored as the same
\texttt{PolicyCallable} on one seed/Wilson-CI/MDE contract
(\S\ref{sec:ladder-contract}). The L1 lift's Wilson intervals
\emph{overlap} ([0.816, 0.866] in common; Cohen's $h{=}0.11$, below the
$N{=}250$ MDE of $0.123$) --- it is reported as within noise, not an
improvement. L2's $\hat{p}{=}0.012$ against a mean rollout coverage near
$0.50$ is the ``learning buys the last fraction of precision'' result.
The L3 rows are an \emph{in-flight} signal, not a finding: the
\emph{same} zero-training world-model planner reaches $9/24{=}0.375$
[0.21, 0.57] on simple navigation and ${\sim}0\%$ on contact manipulation
(\S\ref{sec:l3}, Table~\ref{tab:l3}); the signal is a \emph{single}
DINO-WM family, the Wall Wilson interval spans chance, and the two cells
differ in env, checkpoint, \emph{and} CEM budget (a confounded contrast,
with a \texttt{jepa-wms} second-family de-confound in progress), so we make
no conditional claim.
$^\dagger$auto-downscoped ($n_\text{ep}{=}25$, $N{=}125$;
\S\ref{sec:downscope}). Sources: \texttt{results/ladder/} (L1, L2),
\texttt{results/sweep-full/} (L0, ACT zero-shot), and the cross-repo
\texttt{lerobot-wm-research} (L3).}
\label{tab:ladder}
\end{table}

\paragraph{L0 / L1 --- pretrained and fine-tuned, side by side.}
The pretrained Diffusion Policy reads $\hat{p}{=}0.816$~[0.739, 0.874] on
PushT. On the L1 cell, fine-tuning the already-converged ACT checkpoint on
its own \texttt{aloha\_transfer\_cube} demos moves the success rate from
$0.824$~[0.772, 0.866] ($206/250$) to $0.864$~[0.816, 0.901] ($216/250$).
The point estimate rises $+0.040$, but the two Wilson intervals overlap
on $[0.816, 0.866]$, the paired effect size is small (Cohen's
$h{=}0.11$), and $|\Delta\hat{p}|{=}0.040$ is below the $N{=}250$ MDE of
$0.123$ (\S\ref{sec:downscope}). We therefore report L1 as a
\emph{within-noise, non-significant shift}, not a fine-tuning win. This
is the expected and honest outcome of continued fine-tuning a policy
already near its ceiling on its own training distribution; the negative
result is the finding, and it is kept on the leaderboard's axis rather
than dropped. We do not silently drop the second fine-tune track either:
an attempted SmolVLA LoRA fine-tune \emph{collapsed} the policy
($0.252\to{\sim}0$ on \texttt{libero\_10}\todo{confirm exact post-LoRA
$\hat{p}$ and $N$ from the fine-tune track summary; not committed to this
tree}) rather than lifting it --- a negative we attribute to a small demo
budget against a large adapter and an unfavorable reward structure,
reported as a failed lift, not omitted.

\paragraph{L2 --- a competent controller, and the last 50\% of precision.}
The scripted PushT state-feedback controller is genuinely competent ---
its mean per-episode max-coverage is $0.505$ (median $0.501$),
i.e.\ on an average rollout it pushes the block roughly
halfway to the target footprint --- yet it
clears the strict canonical bar ($\text{coverage}>0.95$) on only
$0.012$~[0.004, 0.035] of episodes ($3/250$). Read against the L0
learned-policy rate on the same env, this is a quantitative statement of
where the value of learning lives: not in getting started, but in the
last fraction of precision a hand-tuned controller cannot close. The
mean-coverage figure is logged in the L2 summary JSON; the binary
$0.012$ is the contract-comparable number.

\subsection{L3 --- an in-flight dynamics-complexity gradient (hypothesis,
not result)}
\label{sec:l3}
This rung is an \emph{in-flight hypothesis} we put on the contract, not a
result. The hypothesis is a \emph{dynamics-complexity gradient}: that
\emph{when} a zero-training world-model planner can substitute for
learning is governed by how modellable the env's dynamics are --- from
point-mass navigation, through contact, to deformable manipulation. This
is the curve we are measuring, not a curve we have measured; the
phrasing throughout is deliberate (we \emph{pose} / \emph{ask} / it is
\emph{in-flight}), and every L3 number below carries its CI, its
single-family caveat, and its cross-env confound in the same sentence. We
take \emph{one} DINO-WM~\cite{zhou2024dinowm} latent-MPC planner --- a
frozen forward model wrapped in a CEM loop, no task-specific training ---
and run it on two envs that probe the ends of that gradient
(Table~\ref{tab:l3}). On \textbf{PushT}, contact-rich manipulation, the
planner scores ${\sim}0$ ($0/2$ honest seeds): it moves the block but
pushes it away from the goal, at ${\sim}1000\times$ the per-step compute
of the learned diffusion policy that scores $0.816$ on the same env. This
is \emph{the contact endpoint of the hypothesized curve} --- a clean data
point where a frozen latent model cannot plan, a feature of the study,
not a failure to rescue with more compute. On \textbf{Wall}, two-room
point-mass navigation, the same planner reaches $9/24{=}0.375$ (Wilson 95\%
[0.212, 0.573], \texttt{max\_steps}${=}40$, success
$:= \lVert\text{goal}-\text{pos}\rVert<4.5$, on an 8\,GB RTX~4060), but the
Wilson interval spans chance and the signal is from a \emph{single}
DINO-WM family, so this is an \emph{in-flight} reading that zero-training
planning can sometimes reach a navigation goal, not a measured success
rate. Critically, Wall and PushT differ in \emph{three} co-varying
dimensions --- env, checkpoint, \emph{and} CEM budget --- so their
cross-env contrast is \emph{confounded} and cannot license the conditional
``planning substitutes for learning on simple dynamics.'' We are running a
second world-model family (\texttt{jepa-wms}) to de-confound the family
axis, but it is throughput-blocked so far. We state the gradient as a
hypothesis the contract is built to test de-confounded (same planner, same
budget, vary only the env, over more than two points), not a finding we
claim here.

\begin{table}[t]
\centering
\small
\begin{tabular}{llrll}
\toprule
Env & Dynamics & $N$ & Success & Per-step compute \\
\midrule
Wall  & 2-room nav (point mass) & 24 & $9/24 = 0.375$ [0.21, 0.57] & \multirow{2}{*}{${\sim}1000\times$ learned policy} \\
PushT & contact manipulation    & 2  & $0/2 \approx 0.0$ & \\
\bottomrule
\end{tabular}
\caption{The L3 in-flight signal (\emph{not} a confirmed conditional): the
\emph{same} zero-training DINO-WM~\cite{zhou2024dinowm} latent-MPC
planner, identical CEM budget ($N{=}300$ samples, $10$ opt.\ iters,
top-$k{=}30$, horizon~$5$), run on two envs. Wall success
$:=\lVert\text{goal}-\text{pos}\rVert<4.5$, \texttt{max\_steps}${=}40$, on
an 8\,GB RTX~4060. The Wall $9/24{=}0.375$ is \emph{indicative only}
($N{=}24$, Wilson 95\% [0.21, 0.57] spans chance, a \emph{single} DINO-WM
family); a fuller-CEM, second-family (\texttt{jepa-wms}) de-confound is
required before any rate is claimed. Upstream
DINO-WM~\cite{zhou2024dinowm} reports near-perfect Wall at H100-scale
$300\times30$ CEM. The Wall-vs-PushT contrast is \emph{confounded} --- env,
checkpoint, and CEM budget all co-vary --- so it is an in-flight signal for
a dynamics-complexity gradient (point-nav $\to$ contact), to be measured
de-confounded, not a direction we confirm here. PushT is the contact
endpoint of that hypothesized curve. Failures are interpretable ---
wall-trap and last-mile stall, both greedy-CEM local minima rather than
bugs. Source: cross-repo \texttt{lerobot-wm-research}, run on the same
eval contract; not an in-embodimetry-tree artifact.}
\label{tab:l3}
\end{table}

A $39\times$ predictor speedup (scaled-dot-product attention,
\texttt{bfloat16}, and encode-once frame caching) is what made any L3
measurement possible on this hardware. The Wall misses are interpretable,
not bugs --- wall-trap (the planner routes into the dividing wall) and
last-mile stall (it arrives near the goal then stalls), both greedy-CEM
local minima a larger budget or longer horizon would be expected to
escape. With $N{=}24$ on Wall (Wilson [0.21, 0.57], spanning chance) and
$N{=}2$ on PushT, a \emph{single} world-model family, and a confounded
cross-env contrast, we make \emph{no} significance claim and \emph{no}
directional claim: what we have is an in-flight signal at two endpoints of
a hypothesized dynamics-complexity gradient. The de-confounded study ---
the same planner at the same budget varying only the env across more than
two points, plus a second world-model family (\texttt{jepa-wms}, currently
throughput-blocked) to break the family confound, at a seed count that
yields a Wilson CI off chance --- is the explicit next step, and until it
runs L3 is in-flight, not a result.

\paragraph{Per-cell leaderboard.}
Table~\ref{tab:leaderboard} gives the v1 learned-policy and floor-baseline
cells on the same Wilson contract, for context around the ladder rungs.

\begin{table}[t]
\centering
\small
\begin{tabular}{llrrl}
\toprule
Policy & Env & $N$ & $\hat{p}$ & 95\% CI \\
\midrule
ACT       & aloha\_xfer    & 250 & 0.824 & [0.772, 0.866] \\
Diffusion & pusht$^\dagger$ & 125 & 0.816 & [0.739, 0.874] \\
SmolVLA   & libero\_goal   & 250 & \textbf{0.928} & [0.889, 0.954] \\
SmolVLA   & libero\_spat.  & 250 & 0.776 & [0.720, 0.823] \\
SmolVLA   & libero\_obj.   & 250 & 0.528 & [0.466, 0.589] \\
SmolVLA   & libero\_10     & 250 & \textbf{0.252} & [0.202, 0.309] \\
random    & aloha\_xfer    & 250 & 0.052 & [0.031, 0.087] \\
random/no-op & other envs  & 2750 & 0.000 & [0.000, 0.002] \\
\bottomrule
\end{tabular}
\caption{v1 per-cell success rate with Wilson 95\% CI, $N{=}5\cdot
n_\text{ep}$ binary outcomes per cell. $^\dagger$auto-downscoped
diffusion$\times$pusht ($n_\text{ep}{=}25$, $N{=}125$;
\S\ref{sec:downscope}). Uniform-random clears floor only on
\texttt{aloha\_transfer\_cube}, so the floor is per-cell. MDE band at
$N{=}250$, $\hat{p}{=}0.5$ is $0.123$; 22~runnable cells; XVLA deferred.}
\label{tab:leaderboard}
\end{table}

\paragraph{Proof the audit bites: a normalization bug we caught in our
own eval.}
A secondary but load-bearing result --- evidence the audit machinery
actually catches errors rather than laundering them --- is a bug in
\emph{our own} harness. ACT's apparent failure on
\texttt{aloha\_transfer\_cube} ($\hat{p}{=}0.016$~[0.006, 0.040], below
the env's random floor) turned out to be ours, not the architecture's. An early
build of the eval loop silently failed to apply dataset normalization stats
to \texttt{observation.images.top}, feeding ACT un-normalized image
observations; PR~\#51 fixed it
(\texttt{\_recover\_dataset\_stats\_from\_safetensors} now disambiguates
buffer names against the config's \texttt{feature\_keys}). After the fix the
same checkpoint reads the canonical \textbf{0.824~[0.772, 0.866]}
(Table~\ref{tab:leaderboard}). To attribute the recovery cleanly we ran a
controlled $2{\times}2$ ablation crossing normalization \{buggy, fixed\}
with inference \{Hub-default, paper-settings\}, each cell $N{=}250$
(Table~\ref{tab:act-ablation}): the recovery is 100\% the normalization
fix, and \emph{temporal ensembling is a wash, not the cause} (broken
normalization stays at 0.016 under either inference setting; once fixed,
Hub-default and paper settings overlap in Wilson CI). The lesson
generalises: an eval suite must be audited --- normalization,
buffer wiring, observation plumbing --- as carefully as the policies it
scores (Section~\ref{sec:results-audit}).

\begin{table}[t]
\centering
\small
\begin{tabular}{lll}
\toprule
& Hub-default inference & Paper-settings inference \\
\midrule
Normalization buggy & 0.016 & 0.016 \\
Normalization fixed & 0.812 & 0.768 \\
\bottomrule
\end{tabular}
\caption{ACT $\times$ \texttt{aloha\_transfer\_cube}, controlled
$2{\times}2$ ablation (normalization $\times$ inference settings), each
cell $N{=}250$ (5~seeds $\times$ 50~episodes). The entire recovery is the
normalization fix (PR~\#51): broken normalization stays at 0.016
regardless of inference settings, and once normalization is fixed the
Hub-default and paper inference settings are statistically
indistinguishable (overlapping Wilson CIs), so temporal ensembling is a
wash. The fixed+Hub-default cell (0.812) is the same condition as the
canonical leaderboard number 0.824~[0.772, 0.866] (Table~\ref{tab:leaderboard})
measured in a separate $N{=}250$ run; the two agree within CI. Source:
\texttt{scripts/probes/probe\_act\_normalization\_ablation.py}.}
\label{tab:act-ablation}
\end{table}

\paragraph{Leaderboard context and baseline sanity.}
For context, SmolVLA's pooled rate across the four LIBERO suites is
0.621~[0.591, 0.651] under our protocol (\texttt{task\_id=0}, env-default
caps), with a sharp per-suite gradient (0.928 \texttt{goal} down to 0.252
\texttt{libero\_10}); these are single-task envelopes, not suite
replications of the paper's 10-task averages~\cite{shukor2025smolvla}
(\S\ref{sec:results-audit}). The floor baselines confirm the bar is real:
no-op succeeds zero times across all six envs ($N{=}1500$, Wilson upper
$0.003$), and uniform-random is above floor only on
\texttt{aloha\_transfer\_cube} ($0.052$~[0.031, 0.087]) --- so the floor
is not one number, which is why we run it on every cell.

\paragraph{Methodology caveats (v1.0.1 audit).}
\label{sec:results-audit}
A static audit surfaced three mismatches that constrain comparative
claims without invalidating any measurement. (1)~\emph{Task scope:} every
SmolVLA cell is \texttt{task\_id=0}, so paper-vs-measured deltas are
single-task envelopes, not replications. (2)~\emph{ACT normalization bug}
(self-caught, PR~\#51): covered above and in Table~\ref{tab:act-ablation};
the $0.016\to0.824$ recovery is entirely the norm fix, temporal ensembling
a wash. (3)~\emph{LIBERO step caps:} v1 inherited lerobot caps shorter than
canonical LIBERO's 600~\cite{liu2023libero}; a cap=600 probe on
\texttt{libero\_10} reads 0.256~[0.206, 0.314] vs.\ 0.252 (within
half-width, policy- not cap-bottlenecked), other suites pending v1.1. No
patch rewrites a parquet row. We rely on label-free failure signals
(episode-length and cap-hit distributions, plus every failed-episode MP4)
rather than a labeled taxonomy, which is specified but deferred to v1.1.
XVLA ran on LIBERO but is deferred from v1 over an unresolved
Hub-artifact wiring bug (\texttt{docs/DEFERRED\_POLICIES.md}).

% ---------------------------------------------------------------- %
% Discussion                                                       %
% ---------------------------------------------------------------- %

\section{Discussion}

\paragraph{What the ladder says.}
Two things are settled and one is in-flight. The two settled, citable
contributions are the cross-paradigm contract and the self-auditing
methodology; the dynamics-complexity gradient is the in-flight hypothesis
the contract was purpose-built to test, and we keep it strictly labeled as
such. The self-audit claim is narrow and defensible: a neutral
seed/Wilson-CI/MDE contract surfaced a normalization bug that pinned
ACT$\times$aloha at $0.016$ (below the random floor) and a clean
$2{\times}2$ ablation attributed the full $0.016\to0.824$ recovery to the
fix (Table~\ref{tab:act-ablation}) --- an error a loss curve hides and only
a closed-loop CI exposes --- and a guard test (\S\ref{sec:guard})
code-enforces that the rung above stays in-flight. At the top of the
ladder we \emph{ask}, but do not answer, \emph{when} a zero-training
world-model planner substitutes for learning. The signal is an
\emph{in-flight} reading at two endpoints of a hypothesized
dynamics-complexity gradient (Wall $9/24{=}0.375$ [0.21, 0.57] vs.\ PushT
${\sim}0$, the contact endpoint of the curve), from a \emph{single}
world-model family, with the Wall Wilson interval spanning chance and the
cross-env contrast confounded by co-varying env, checkpoint, and CEM
budget --- so we hold it as an \emph{in-flight hypothesis}, not a result,
and the contract is what makes the de-confounded version (a second
\texttt{jepa-wms} family included) measurable.

The supporting spine is a set of honest negatives that earn the audit its
credibility. First, the contract works as a cross-paradigm yardstick:
pretrained, fine-tuned, classical, and world-model policies ran through
one loop, were seeded identically, and are read off the same Wilson
interval (Tables~\ref{tab:ladder},~\ref{tab:l3}); every L0--L2 row is
bit-replayable and Wilson/bootstrap CIs agree within one half-width.
Second, the measured rungs say something specific and undramatic:
fine-tuning a near-ceiling policy moves it within noise ($+0.040$,
overlapping CIs, $h{=}0.11$) and an attempted SmolVLA LoRA fine-tune
collapses rather than lifts, so neither L1 track is inflated into a win;
and a competent classical controller gets halfway to the goal on average
but clears the strict bar on ${\sim}1\%$ of episodes ($\hat{p}{=}0.012$),
locating the value of learning in the last fraction of precision. Third,
the machinery catches its own errors: ACT$\times$aloha's $0.016\to0.824$
recovery (Table~\ref{tab:act-ablation}) was a normalization bug in the
harness, not the architecture. The two contributions we stand behind
without qualification are (i)~the \emph{contract} --- one
seed/Wilson-CI/MDE spine from pretrained policies to world-model planners,
with floor baselines and honest negatives retained, sub-MDE deltas flagged
inconclusive, and overlapping CIs called overlapping --- and (ii)~the
\emph{self-audit} --- the instrument correcting its own $0.016\to0.824$
overclaim, with a guard test (\S\ref{sec:guard}) code-enforcing that the
top rung stays in-flight. The dynamics-complexity gradient is neither: it
is the in-flight hypothesis the contract was purpose-built to test, kept
strictly labeled --- a confounded, single-family, top-of-ladder signal
(Wall $9/24{=}0.375$ [0.21, 0.57] vs.\ the PushT contact endpoint) we
\emph{pose}, not a finding --- and it earns its own citation only once the
de-confound (second \texttt{jepa-wms} family, canonical-budget rerun)
lands in v1.1.

\subsection{The L3 hypothesis and what would confirm it}
The dynamics-complexity gradient is a hypothesis, not a conditional we can
yet assert, and it is worth stating cleanly because the contract is built
to test it. The intuition: Wall's point-mass dynamics are
low-dimensional, near-linear, and contact-free, so a frozen latent
forward model may predict them accurately enough that greedy CEM finds a
goal-reaching plan it was never trained on; PushT's dynamics are
contact-rich and stiff, so small latent errors compound through contact
and the planner pushes the block confidently the wrong way. If that
intuition holds, ``do we need a learned policy?'' becomes a question about
the modellability of the dynamics rather than about the planner. But our
current evidence cannot establish it: the Wall cell is $N{=}24$ with a
Wilson interval [0.21, 0.57] that spans chance, the signal is from a
\emph{single} DINO-WM family, and Wall and PushT differ in env,
checkpoint, \emph{and} CEM budget, so the apparent nav-vs-manip
\emph{direction} is confounded by all three. What would confirm the
gradient is the de-confounded study the contract enables: the \emph{same}
planner at the \emph{same} budget across a \emph{sequence} of envs
(point-nav $\to$ contact $\to$ deformable), each at a seed count whose
Wilson CI clears chance, plus a second world-model family
(\texttt{jepa-wms}) to break the single-family confound, with the
planner's compute cost (${\sim}1000\times$ per step) reported alongside.
Until then the defensible statement is narrow: a zero-training planner can
\emph{sometimes} reach a navigation goal it was never trained on
(in-flight), and fails the PushT contact endpoint of the hypothesized
curve --- two points, not a curve, and not a substitution claim.

\subsection{Limitations}

\begin{itemize}\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}\setlength{\topsep}{0pt}\setlength{\partopsep}{0pt}
    \item \textbf{Simulation only.} All numbers are sim; the sim-to-real
    gap is unmeasured. embodimetry is a screening tool, not a
    substitute for hardware evaluation.
    \item \textbf{Single-hardware wall-clock.} Per-step latency is from
    one RTX~4060 laptop; absolute ms/step differs across GPUs, but the
    \emph{ranking} the auto-downscope rule consumes is portable.
    \item \textbf{$\boldsymbol{\pi_0}$ family deferred.} $\pi_0$,
    $\pi_0$-FAST, $\pi_{0.5}$ (PaliGemma-3B) need ${\approx}30$\,GB host
    RAM during \texttt{from\_pretrained} cold load ($<4$\,GB peak VRAM),
    exceeding the 32\,GB v1 laptop's headroom; v1 compares four learned
    policies, not seven (v1.1 uses quantized/streaming load).
    \item \textbf{L3 is an in-flight, confounded, single-family signal ---
    not a result.} We make no significance and no directional claim at the
    top of the ladder. The Wall cell is $N{=}24$ seeds with a Wilson
    interval [0.21, 0.57] wide enough to span chance, and PushT is $N{=}2$.
    \emph{The cross-env confound is itself an open limitation we state
    plainly:} the Wall-vs-PushT contrast varies \emph{three} dimensions at
    once --- (i)~the \emph{env} (point-nav vs.\ contact manipulation),
    (ii)~the \emph{checkpoint} (a different DINO-WM forward model per env),
    and (iii)~the \emph{CEM budget} relative to each env's difficulty ---
    so any apparent ``simple dynamics $\to$ planning works'' direction is
    confounded by checkpoint and budget, not isolated to dynamics; and the
    whole signal comes from a \emph{single} world-model family. The budget
    itself is reduced to fit 8\,GB ($N{=}300$/opt~$10$/top-$k$~$30$/
    horizon~$5$, vs.\ the source method's H100-scale $300\times30$). We
    therefore frame L3 as an in-flight signal for a dynamics-complexity
    gradient \emph{hypothesis}, to be confirmed only by a de-confounded
    study (same planner and budget, vary only the env, $\geq3$ points, a
    second \texttt{jepa-wms} family, seed count clearing chance); that
    second-family de-confound is in progress but throughput-blocked. The L3
    cells are also a \emph{cross-repo} artifact
    (\texttt{lerobot-wm-research}), not an in-tree parquet of this
    repository.
    \item \textbf{Single-GPU compute ceiling.} The entire audit runs on
    one 8\,GB RTX~4060 laptop. This caps the ladder beyond L3: SmolVLA
    full fine-tune does not fit, so L1 is a single ACT cell plus an
    attempted SmolVLA LoRA track that collapsed ($0.252\to{\sim}0$),
    reported as a failed lift rather than dropped.
    \item \textbf{L1 is one cell, and the lift is within noise.} The
    fine-tuning rung is ACT$\times$aloha alone, and its $+0.040$ shift has
    overlapping Wilson CIs and $h{=}0.11$ (below the $N{=}250$ MDE). We
    cannot claim fine-tuning helps from this cell; we can only show the
    contract reads a near-ceiling continued-FT as the non-result it is. A
    multi-cell L1 sweep with a larger demo budget and a from-scratch
    baseline is needed to say anything stronger.
    \item \textbf{SmolVLA single-task scope.} Every SmolVLA cell is
    \texttt{task\_id=0} only; the paper~\cite{shukor2025smolvla} averages
    10~tasks $\times$ 10~trials. Our SmolVLA numbers are single-task
    envelopes, not suite replications, and the cross-paradigm comparison
    should not lean on them.
    \item \textbf{Sparse matrix, multi-comparison, and episode budget.}
    Cross-policy comparisons are valid only within shared envs; the
    L0/L1/L2 rungs sit on \emph{different} envs (PushT, aloha), so they are
    same-contract but not same-task and must be read as paradigm
    snapshots, not a head-to-head. The L3 in-flight signal is the exception
    that motivates the design: it holds the \emph{policy} fixed and varies the
    \emph{env}, which is the cleanest cross-env comparison the contract
    supports --- but it is only two envs, so ``simple vs.\ hard dynamics''
    is a two-point contrast, not a curve. With several pairwise claims in
    play we control family-wise error by Holm--Bonferroni, but the matrix
    is too sparse for a fully crossed comparison. $N{=}250$ is a
    GPU-budget choice; claims cite only deltas above the per-cell MDE.
    \item \textbf{Harness bugs and protocol-vs-paper.} Our own
    normalization bug (PR~\#51) shows a harness can be the dominant error
    source; the LIBERO cap mismatches make those cells protocol-vs-paper,
    not replication failures. Cross-policy paired tests defer to v1.1 when
    XVLA returns.
\end{itemize}

\paragraph{Future work.}
De-confound and measure the L3 hypothesis: rerun the \emph{same} DINO-WM
planner at a larger seed count and a fuller (paper-faithful) CEM budget on
a GPU that affords it, holding the planner and budget fixed while varying
\emph{only} the env across $\geq3$ points (point-nav $\to$ contact $\to$
deformable), so the indicative Wall $9/24{=}0.375$ becomes a CI-bearing
rate off chance and ``simple vs.\ hard dynamics'' becomes a measured
gradient rather than a confounded, single-family two-point signal; add a
second world-model family (\texttt{jepa-wms}, with
V-JEPA-2-AC~\cite{assran2025vjepa2}) to break the family confound. The
gradient earns its own citation in v1.1 once that de-confound and the
canonical-600 rerun land. Broaden L1 to a
multi-cell, larger-budget fine-tuning sweep with a from-scratch baseline
so the rung can support a positive claim, and diagnose the SmolVLA LoRA
collapse. Close the audit caveats (cap=600 LIBERO reruns); onboard the
$\pi_0$ family via quantized/streaming load; and re-run on hardware for
the sim-to-real gap.

% ---------------------------------------------------------------- %
% References                                                       %
% ---------------------------------------------------------------- %

{\footnotesize
\setlength{\itemsep}{0pt}
\bibliographystyle{plain}
\bibliography{references}
}

\end{document}
